---
title: "HW 2"
author: "Hongzhu Ren"
date: "2023-10-01"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1
```{r pol_month, include=FALSE}
# import and clean pol_month
pol_month_clean <- read_csv("./data/pols-month.csv") |> 
  separate(
    col = mon,
    into = c("year","month","day"),
    sep = "-"
  ) |> 
  mutate(
    across(c("year","month","day"),as.numeric),
    month = as.character(factor(month, level = 1:12, month.abb)),
    president = ifelse(prez_gop == 1, "gop", "dem")
  ) |> 
  select(
    -prez_dem,
    -prez_gop,
    -day
  )
```

```{r,echo=FALSE}
pol_month_clean
```

**pol_month** data described the party composition of government agencies ranging from `r pol_month_clean$year[1]` to `r pol_month_clean$year[length(pol_month_clean$year)]`. The prefix *gov*,*sen*,*rep* meant separately the number of governors, the number of senators,the number of representatives. And the suffix *gop* and *dem* referred to the two parties of USA.

```{r snp, include=FALSE}
# import and clean snp
snp_clean <- read_csv("./data/snp.csv") |> 
  separate(
    col = date,
    into = c("month","day","year"),
    sep = "/"
  ) |> 
  mutate(
    across(c("year","month","day"),as.numeric),
    year = ifelse(year < 23,year+2000,year+1900),
    month = factor(month, level = 1:12, month.abb)
  ) |> 
  select(
    -day,
  ) |> 
  select(
    "year","month",everything()
  ) |> 
  arrange(
    year,month
  ) 
```

```{r,echo=FALSE}
snp_clean
```
**snp** data reflected the stock performance . The *close* variable referred to the closing values of the S&P stock index on the associated date ranging from `r snp_clean$year[1]` to `r snp_clean$year[length(pol_month_clean$year)]`

```{r unemployment,include=FALSE}
# import and clean unemployment 
unemployment_clean <- read_csv("./data/unemployment.csv") |> 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  ) |> 
  select(
    year = Year,everything()
  )

#pol_month_clean_m <- mutate(
  #pol_month_clean,
  #date <- ymd(paste(pol_month_clean$year, match(pol_month_clean$month,month.abb)))
#)
```

```{r,echo=FALSE}
unemployment_clean
```
**unemployment** data described the unemployment rate over time.

```{r merge,echo=FALSE}
#merge data   
data_merge <- merge(pol_month_clean,snp_clean,all = TRUE)
data_merge <- merge(data_merge,unemployment_clean,all = TRUE)
data_merge <- tibble(data_merge)
data_merge
```

 The final data is a `r ncol(data_merge)` columns and `r nrow(data_merge)` rows dataset. This data showed government composition in association with stock performance and unemployment situation ranging from `r data_merge$year[1]` to `r data_merge$year[length(data_merge$year)]`.


# Problem 2 
import and clean Mr.Trash Wheel data
```{r trash wheel,echo=FALSE}
trash <- read_excel("./data/202207 Trash Wheel Collection Data.xlsx",
                    sheet = 1,
                    range = cell_cols("A:N")) |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(
    homes_powered = weight_tons*500/30
  )

print(trash)
```
 
import and clean Professor and Gwynnda data
```{r professor&gwynnda,include=FALSE}
professor <- read_excel("./data/202207 Trash Wheel Collection Data.xlsx",
                    sheet = 2,
                    range = cell_cols("A:M")) |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(
    homes_powered = weight_tons*500/30
  )

gwynnda <- read_excel("./data/202207 Trash Wheel Collection Data.xlsx",
                    sheet = 4,
                    range = cell_cols("A:K")) |> 
  janitor::clean_names() |> 
  drop_na(dumpster) |> 
  mutate(
    homes_powered = weight_tons*500/30
  )
```

Add tag to each dataset

```{r tag, include=FALSE}
trash_comb <- trash |> 
  mutate(
    tag = "Mr.trash"
  ) |> 
  select(tag,everything())

professor_comb <- professor |> 
  mutate(
    tag = "professor"
  ) |> 
  select(tag,everything())

gwynnda_comb <- gwynnda |> 
  mutate(
    tag = "gwynnda"
  ) |> 
  select(tag,everything())
```

```{r}
trash_comb <- mutate(trash_comb,year = as.numeric(year))
trash_final <- bind_rows(trash_comb,professor_comb,gwynnda_comb)

```
The final combined data is a `r ncol(trash_final)` and `r nrow(trash_final)`, the *weight_tons* variable refers to the total weight the trash wheel carried each day with the unit ton. *home_powered* estimate the number of houses which can be supplied by energy generated from trash wheel.

As for the available data, the total weight of trash collected by Professor Trash Wheel was `r sum(trash_final[trash_final$tag=="professor",]$weight_tons)` tons; the total cigarette butts collected by Gwynnda in July of 2021 was `r sum(filter(gwynnda,month=="July",year==2021)$cigarette_butts)`. 

#problem 3